curling

### 2020-07-30 inst/f1points.Rmd
more work

---
documentclass: jss
author:
  - name: Robin K. S. Hankin
    affiliation: Auckland University of Technology
    address: >
      2-14 Wakefield Street
      Auckland, New Zealand
    email: \email{hankin.robin@gmail.com}
    url: http://aut.ac.nz/
title:
  formatted: "An Objective Assessment of Points Scorings in Formula 1 Motor Racing"
  plain:     "An Objective Assessment of Points Scorings in Formula 1 Motor Racing"
  # For running headers, if needed
  short:     "Formula 1 Points Scoring"
abstract: >
  The point scoring system of Formula 1 motor racing is a long-standing
  contention among fans and competitors, with changes to the points
  scored occurring frequently.  Here, I apply generalized Plackett-Luce
  likelihood functions to Formula 1 racing, and assess historical points
  scoring systems objectively.  In a well-defined sense, the points
  scoring system of 1985 is the best of the historical ones.  I make
  similar observations about related competitive situations including
  the Eurovision Song contest and XXX.
keywords:
  # at least one keyword must be supplied
  formatted: [keywords, not capitalized, "\\proglang{Java}"]
  plain:     [keywords, not capitalized, Java]
preamble: 
  \usepackage{amsmath}
output: rticles::jss_article
---

@@

We now assess the six points systems using the methodology outlined
above, using 1000 _in silico_ trials.  For each of the six points
systems, each of the 1000 trials results in a single non-negative
integer: the degree of agreement between the Plackett-Luce ranking and
the rankings according to the points system considered.  There are
three measures that might be used to assess the distribution of degree
of agreement: (1), the mean degree of agreement $\overline{d}$; (2),
the probability of correctly predicting the winner,
$\operatorname{Pr}(d\geqslant 1)$; and (3), the probability of
correctly predicting the complete order statistic
$\operatorname{Pr}(D=n)$.  These summaries are shown for each of the
six points systems in figures \ref{simulated_mean},
\ref{simulated_winner_correct} and \ref{simulated_complete_agreement}
respectively.

@@

Points systems are a source of lively debate from many perspectives.
The point of a point allocation system is to rank competitors in an
objective way; but

Here I treat total points scored as a random variable and compare
different point allocation schemes against both one another and

but until now there is no objective way of comparing points systems.
if one considers order tables to be a random variable.  Because the
numerical value of points awarded for first, second, third etc, does
not affect driver behaviour it is a reasonable requirement that

Here, I assess
several plausible points systems including the one actually used in
Formula 1.  By several criteria, the

>>>>

---
documentclass: jss
author:
  - name: Robin K. S. Hankin
    affiliation: Auckland University of Technology
    address:
      2-14 Wakefield Street
      Auckland, New Zealand
    email: \email{hankin.robin@gmail.com}
    url: http://aut.ac.nz/
title:
  formatted: "An Objective Assessment of Points Scorings in Formula 1 Motor Racing"
  plain:     "An Objective Assessment of Points Scorings in Formula 1 Motor Racing"
  # For running headers, if needed
  short:     "Formula 1 Points Scoring"
abstract: >
  The points scoring system of Formula 1 motor racing is a long-standing
  contention among fans and competitors, with changes being controversial.
  The inaugural points system, in use from 1950
  to 1959 was $(8,6,4,3,2)$: 8 points to the winner, 6 to second
  place, and so on.  In 2020 the points system was changed to
  $(25,18,15,12,10,8,6,4,2,1)$.  The overall ranking of the
  competitors is determined by summing points awarded over a season;
  but this system suffers from the defect that the points system
  itself is intrinsically arbitrary.  Here, I apply generalized
  Plackett-Luce likelihood functions to Formula 1 motor racing, and
  use maximum likelihood techniques to identify a ranking for the competitors;
  this method includes no use of an arbitrary points system.
  Using this methodology, I assess a number of reasonable points
  scoring systems objectively.  In a well-defined statistical sense, a
  ``winner-takes-all'' points system outperforms the other points
  systems considered.
keywords: 
  # at least one keyword must be supplied
  formatted: [Bradley-Terry, Plackett-Luce, order statistics, rank
    statistics, points systems, Formula 1]
  plain:     [Bradley-Terry, Plackett-Luce, order statistics, rank
    statistics, points systems, Formula 1]
preamble:
  \usepackage{amsmath}
  \usepackage{amssymb}
output: rticles::jss_article
---

@@

We now assess the six points systems using the methodology outlined
above, using 1000 _in silico_ trials.  For each of the six points
systems, each of the 1000 trials results in a single non-negative
integer: the degree of agreement between the Plackett-Luce ranking and
the rankings according to the points system considered.  There are
three measures that might be used to assess the distribution of degree
of agreement: (1), the mean degree of agreement $\overline{d}$; (2),
the probability of correctly predicting the winner,
$\operatorname{Pr}(d\geqslant 1)$; and (3), the probability of
correctly predicting the complete order statistic
$\operatorname{Pr}(D=n)$.  These summaries are shown for each of the
six points systems in figures \ref{simulated_mean},
\ref{simulated_winner_correct} and \ref{simulated_complete_agreement}
respectively.  These figures differ in detail but all have one
feature: the winner-takes-all points system consistently outperforms
the other systems.  This is particularly evident in figure
\ref{simulated_complete_agreement}.

@@

Many competitive situations involve ranking competitors and one common
method for combining repeated observations is to assign points for the
winner, second place, third place, etc.  The overall ranking is
decided on the basis of accumulated points.  However, statistical
analysis made difficult by the arbitrary values of the points given.

The points system used in Formula 1 motor racing is a source of lively
debate from many perspectives, with changes being controversial.  
Here I treat total points scored as a random variable and compare
different point allocation schemes against objective Plackett-Luce
ranks.  By several criteria, a ``winner takes all'' performs better
than other reasonable points systems.  However, such a system
furnishes little incentive for mid-ranking competitors and as such
might result in duller races.
